%%
%% This is file `thesis-ex.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% uiucthesis2014.dtx  (with options: `example')
%% 
\def\fileversion{v2.25b} \def\filedate{2017/05/02}
%% Package and Class "uiucthesis2014" for use with LaTeX2e.
\documentclass[12pt,edeposit,fullpage]{uiucthesis2014}

% references at the end of each chapter
\usepackage[round,sectionbib]{natbib}
\usepackage{chapterbib}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bibentry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{rotating}
\usepackage{listings}
\usepackage{fancyvrb}

\usepackage{lscape}
\usepackage{longtable}


\usepackage{adjustbox}


\usepackage[ruled,vlined,linesnumbered]{algorithm2e}

\usepackage{listings} % code in snippets
\usepackage{color} % color syntax

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}
\usepackage[T1]{fontenc}

\DeclareRobustCommand{\firstsecond}[2]{#1}

\usepackage[titletoc]{appendix}

\lstset{ %
  backgroundcolor=\color{white},   % choose the background color
  basicstyle=\footnotesize,        % size of fonts used for the code
 frame=single,	   
  breaklines=true,                 % automatic line breaking only at whitespace
    numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,
  numberstyle=\color{mygray},
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  keywordstyle=\color{blue},       % keyword style
  stringstyle=\color{mymauve},     % string literal style
    showspaces=false,
  showstringspaces=false,
  upquote=true,
}
%  literate={``}{\textquotedblleft}1,  %set right quote mark
 % deletestring=[b]",  % del left quote mark
  
\usepackage{multirow}
\usepackage[normalem]{ulem}
\useunder{\uline}{\ul}{}

% First level headings in table of contents
\setcounter{tocdepth}{1}


\def\aap{{A\&A}}
\def\AaA{{A\&A}}
\def\AJ{{AJ}}
\def\aj{{AJ}}
\def\ApJ{{ApJ}}
\def\apj{{ApJ}}
\def\apjl{{ApJL}}
\def\ApJL{{ApJL}}
\def\ApJSupp{{ApJS}}
\def\MN{{MNRAS}}
\def\mnras{{MNRAS}}
\def\mnrasl{{MNRAS Letters}}
\def\pasp{{PASP}}
\def\jcap{JCAP}
\def\pasj{{PASJ}}
\def\nat{Nature}
\def\ARAA{{ARA\&A}}
\def\apjs{{ApJS}}
\def\araa{{ARA\&A}}
\def\apss{{Ap\&SS}}
\def\cjaa{{ChJAA}}
\def\aaps{{A\&AS}}

\newcommand{\eg}{{e.g., }}
\newcommand{\ie}{{i.e., }}
\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

%Define Robert Comment

\newcommand\RC[1]{\color{red}{\textbf{ Robert Comment: }{#1} }\color{black}}

% put figure on a new page
\renewcommand{\floatpagefraction}{0.1}

\newsavebox{\FVerbBox}
\newenvironment{FVerbatim}
 {\VerbatimEnvironment
  \begin{center}
  \begin{lrbox}{\FVerbBox}
  \begin{BVerbatim}}
 {\end{BVerbatim}
  \end{lrbox}
  \fbox{\usebox{\FVerbBox}}
  \end{center}}



\makeatletter
\newenvironment{breakablealgorithm}
  {% \begin{breakablealgorithm}
   \begin{center}
     \refstepcounter{algorithm}% New algorithm
     \hrule height.8pt depth0pt \kern2pt% \@fs@pre for \@fs@ruled
     \renewcommand{\caption}[2][\relax]{% Make a new \caption
       {\raggedright\textbf{\ALG@name~\thealgorithm} ##2\par}%
       \ifx\relax##1\relax % #1 is \relax
         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##2}%
       \else % #1 is not \relax
         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##1}%
       \fi
       \kern2pt\hrule\kern2pt
     }
  }{% \end{breakablealgorithm}
     \kern2pt\hrule\relax% \@fs@post for \@fs@ruled
   \end{center}
  }
\makeatother

\nocopyrightpage

\begin{document}

\title{Machine Learning and Information-Theoretic Approaches for Financial Applications}
\author{Kelechi Moise Ikegwu}
\department{Informatics}
\phdthesis
\advisor{Robert J.\ Brunner}
\degreeyear{2021}
\committee{
    Professor Robert J.\ Brunner,  Chair,  Director of Research\\
    Professor Theodore Sougiannis\\
    Assistant Professor Jeff McMullin\\
    Assistant Professor Joseph Yun}
\maketitle

\frontmatter

%% Create an abstract that can also be used for the ProQuest abstract.
%% Note that ProQuest truncates their abstracts at 350 words.
\begin{abstract}


Firm-specific information contributes to resource allocation in capital markets.  The prediction and inference of future firm-specific information are often critical factors that can lead to a prosperous economy.  Machine Learning and Information Theory offer new, alternative methods to aid in the prediction and inference of future firm-specific information.  Thus, we explore approaches from machine learning and information theory to improve existing research areas in accounting and finance.  
In Chapter \ref{Chapter:ABIS}, we utilized random forest trees to predict the annual direction of profitability for firms with a minimal amount of information.  We demonstrate that we can out-perform benchmark models typically used in financial accounting research.  
%Maybe Remove SML?
%In Chapter \ref{Chapter:SML}, we introduce a programming language agnostic framework called SML that integrates a query-like language to simplify the development of machine learning pipelines.  We show that SML significantly reduces the code one has to write to develop the constituent parts of a machine learning pipeline. 


Major public firms announce their annual earnings during the first quarter of the year.  The embedding information in these announcements effects the announcing firm and is transferred to or incorporated by other firms.  Existing literature in finance and accounting focuses on measuring information transfers as effects stemming from firm-specific information releases.  In Chapter \ref{Chapter:PyIF}, we discuss a solution to estimate information transfer via transfer entropy between random processes.  We show that our solution to estimate information transfer scales better with data set size and is up to 1,072 times faster than all existing, open source solutions for large datasets.  In Chapter \ref{Chapter:EATE}, we present an alternative approach to estimate information transfer between firms centered around earnings announcements.  We show that information transfer between firms is stronger for firms on days with unexpected earnings news.  We also demonstrate that communities of firms significantly differ depending on the presence of earnings releases. 

\end{abstract}

%% Create a dedication in italics with no heading, centered vertically
%% on the page.
\begin{dedication}
To my Father,  Mother,  Joshua,  Sandra,  Ezinne,  Natalia,  Khamille,  and Khalil.
\end{dedication}

%% Create an Acknowledgements page, many departments require you to
%% include funding support in this.
\chapter*{Acknowledgments}

%\RC{Kelechi, the pleasure has been all mine. You have been a fantastic student and I have enjoyed working with you and, more importantly, getting to know you as a person. I hope we can stay in touch, and I look forward to your continued success!}

First and foremost, I would like to thank my advisor,  Robert Brunner, for his guidance,  patience,  and support during my graduate studies at the University of Illinois at Urbana-Champaign.
I also wish to thank my committee members: Theodore Sougiannis, Jeff McMullin, and Joseph Yun,  for their help and support.
I am also grateful to Junhyung Kim, Vincent Reverdy, William Biscarri,  Samantha Thrush,  Matias Carrasco Kind,  Jacob Trauger,  Alice Perng, and Howard Hardiman.

\begin{flushright}
--- Kelechi Moise Ikegwu
\end{flushright}

\vspace{10mm}

This work was partially funded by the Graduate College Fellowship program at the University of Illinois.  The work was partially funded by the National GEM Consortium.  This work utilizes resources provided by the Innovative Systems Laboratory at the National Center for Supercomputing Applications at the University of Illinois at Urbana-Champaign.

%% The thesis format requires the Table of Contents to come
%% before any other major sections, all of these sections after
%% the Table of Contents must be listed therein (i.e., use \chapter,
%% not \chapter*).  Common sections to have between the Table of
%% Contents and the main text are:
%%
%% List of Tables
%% List of Figures
%% List Symbols and/or Abbreviations
%% etc.

\tableofcontents
%\listoftables
%\listoffigures


    

%% Create a List of Abbreviations. The left column
%% is 1 inch wide and left-justified
%\chapter{List of Abbreviations}
%
%\begin{symbollist*}
%\item[CA] Caffeine Addict.
%\item[CD] Coffee Drinker.
%\end{symbollist*}

%% Create a List of Symbols. The left column
%% is 0.7 inch wide and centered
%\chapter{List of Symbols}
%
%\begin{symbollist}[0.7in]
%\item[$\tau$] Time taken to drink one cup of coffee.
%\item[$\mu$g] Micrograms (of caffeine, generally).
%\end{symbollist}

\mainmatter

\include{1-introduction}
%\include{2-SML}
\include{3-PredProfitML}
\include{4-PyIF}
\include{5-EarningsAnnouncementsInfoTransfers}
%\include{6-CommunityDetection}
\include{7-Conclusion}




\chapter*{References}
\renewcommand{\bibsection}{}
\bibliographystyle{plainnat}
\bibliography{thesisbib}


%\appendix
%\input{Appendix.tex}
%\include{Appendix}

\backmatter

\end{document}
\endinput
%%
